Evaluation of performances

### 2019-12-09 analysis/eval-performance.Rmd
add scatterplots to vis BM perf

```{r, warning=FALSE}
df_simple = getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
  mutate(task.id = recode_factor(task.id,
    `defoliation-all-plots-HR` = "HR",
    `defoliation-all-plots-VI` = "VI",
    `defoliation-all-plots-NRI` = "NRI",
    `defoliation-all-plots-HR-NRI` = "HR-NRI",
    `defoliation-all-plots-HR-VI` = "HR-VI",
    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
  )) %>%
    tidyr::separate(learner.id, c("learner_group", "filter"),
    remove = FALSE,
    sep = " "
  )
```

@@

```{r eval-performance-1, warning=FALSE}
table1 <- df_simple %>%
  group_by(learner_group, task.id, filter) %>%
  summarise(best = min(rmse.test.rmse)) %>%
  dplyr::rename(
    "Learner Group" = learner_group,
    "Task" = task.id,
    "Filter" = filter,
    "RMSE" = best
  ) %>%
  mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
  arrange(RMSE)

@@

table1 %>%
  flextable()
```

@@

*-CV: L2 penalized regression using the internal 10-fold CV tuning of the `glmnet` package
*-MBO: L2 penalized regression using using MBO for hyperparameter optimization.

@@

```{r performance-results, warning=FALSE, dev = c("png", "pdf")}
results_aggr <- df_simple %>%
  mutate(filter = replace(filter, is.na(filter), "NF")) %>%
  mutate(learner_group = recode_factor(learner_group, `XG` = "XGBOOST")) %>%
  group_by(learner_group, task.id) %>%
  # get the best performance per learner and task
  # this is better than summarise() because we can keep additional columns
  # in constrast to summariuse which only keeps the grouping columns and the
  # summarised one
  slice(which.min(rmse.test.rmse))

@@

```{r filter-effect, warning=FALSE, dev = c("png", "pdf")}
# we rbind two filtered data.frames
# doing the filtering in one step is (at least) complicated here

best_filter <- df_simple %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter != "No Filter") %>%
  group_by(learner_group, task.id) %>%
  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
  slice(which.min(rmse.test.rmse))

@@

best_filter <- df_simple %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter != "No Filter") %>%
  group_by(learner_group, task.id) %>%
  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
  slice(which.min(rmse.test.rmse))

borda <- df_simple %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter == "Borda") %>%
  group_by(learner_group, task.id) %>%
  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
  slice(which.min(rmse.test.rmse))

@@

```{r filter-perf-all, warning=FALSE, dev = c("png", "pdf")}

perf_all_filters <- df_simple %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter != "No Filter") %>%
  group_by(learner_group, task.id) %>%
  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost")

@@

>>>>

```{r}
df_perf = getBMRPerformances(bm_aggregated, as.df = T) %>%
  mutate(task.id = recode_factor(task.id,
    `defoliation-all-plots-HR` = "HR",
    `defoliation-all-plots-VI` = "VI",
    `defoliation-all-plots-NRI` = "NRI",
    `defoliation-all-plots-HR-NRI` = "HR-NRI",
    `defoliation-all-plots-HR-VI` = "HR-VI",
    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
  )) %>%
  tidyr::separate(learner.id, c("learner_group", "filter"),
    remove = FALSE,
    sep = " "
  ) 
```

Aggregate performances and add standard error column.

```{r aggr-perf}
df_perf %<>% 
  group_by(task.id, learner.id, filter) %>% 
  mutate(rmse_aggr = round(mean(rmse), 3)) %>% 
  mutate(se = round(sd(rmse), 3)) %>% 
  select(-rmse, iter) %>% 
  ungroup()
```

@@

```{r eval-performance-1, warning=FALSE}
table1 <- df_perf %>%
  group_by(learner.id, task.id, filter) %>%
  slice(which.min(rmse_aggr)) %>%
  dplyr::rename(
    "Model" = learner_group,
    "Learner ID" = learner.id,
    "Task" = task.id,
    "Filter" = filter,
    "RMSE" = rmse_aggr,
    "Standard Error" = se,
  ) %>%
  ungroup() %>% 
  mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
  select(-iter, -`Learner ID`, -rsq, -expvar) %>% 
  arrange(RMSE)

@@

table1 %>%
  flextable() %>% 
  autofit()
```

@@

*CV: L2 penalized regression using the internal 10-fold CV tuning of the `glmnet` package

*MBO: L2 penalized regression using using MBO for hyperparameter optimization.

@@

```{r performance-results, warning=FALSE, dev = c("png", "pdf")}
results_aggr <- df_perf %>%
  mutate(filter = replace(filter, is.na(filter), "NF")) %>%
  mutate(learner_group = recode_factor(learner_group, `XG` = "XGBOOST")) %>%
  group_by(learner_group, task.id) %>%
  # get the best performance per learner and task
  # this is better than summarise() because we can keep additional columns
  # in constrast to summariuse which only keeps the grouping columns and the
  # summarised one
  slice(which.min(rmse.test.rmse))

@@

```{r filter-effect-best-vs-no-filter, warning=FALSE, dev = c("png", "pdf")}
# we rbind two filtered data.frames
# doing the filtering in one step is (at least) complicated here

best_filter <- df_perf %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter != "No Filter") %>%
  group_by(learner_group, task.id) %>%
  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
  slice(which.min(rmse.test.rmse))

@@

best_filter <- df_perf %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter != "No Filter") %>%
  group_by(learner_group, task.id) %>%
  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
  slice(which.min(rmse.test.rmse))

borda <- df_perf %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter == "Borda") %>%
  group_by(learner_group, task.id) %>%
  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
  slice(which.min(rmse.test.rmse))

@@

```{r filter-perf-all, warning=FALSE, dev = c("png", "pdf")}
perf_all_filters <- df_perf %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter != "No Filter") %>%
  group_by(learner_group, task.id) %>%
  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost")

@@

## (Plot) Scatterplots of filter methods vs. no filter for each learner and task

Showing the final effect of applying feature selection to a learner for each task.
All filters are summarized into a boxplot whereas using "no filter" appears as a standalone observation.

```{r filter-effect-all-vs-no-filter, warning=FALSE, dev = c("png", "pdf")}
results_aggr1 <- df_perf %>%
  filter(learner_group != "Ridge-MBO", learner_group != "Ridge-CV") %>%
  filter(learner_group != "Lasso-MBO", learner_group != "Lasso-CV") %>%
  mutate(filter = replace(filter, is.na(filter), "NF")) %>%
  mutate(learner_group = as.factor(learner_group)) %>%
  mutate(learner_group = recode(learner_group, `XGBoost` = "XG")) %>%
  group_by(learner_group, task.id, filter) %>% 
  # we actually took the mean already in chunk 'aggr-perf'. This is only to get
  # summarise() working
  summarise(perf = mean(rmse_aggr))

results_aggr1 %>%
  ggplot(aes(x = perf, y = learner_group)) +
  geom_beeswarm(
    data = results_aggr1[results_aggr1$filter != "NF", ], size = 1.1, shape = 3,
    groupOnX = FALSE, aes(color = "Filter")
  ) +
  geom_point(data = results_aggr1[results_aggr1$filter == "NF", ], 
             size = 2, shape = 19, aes(color = "No Filter")) +
  facet_wrap(~task.id) +
  scale_color_nejm() +
  labs(x = "RMSE", y = "Task", colour = NULL) +
  guides(size = FALSE) +
  scale_x_continuous(limits = c(30, 50)) +
  theme_pubr() +
  theme(
    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
    axis.title.y = element_blank(),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
  )
```

## (Plot) Scatterplots of filter methods vs. Borda for each learner and task

Showing the final effect of applying feature selection to a learner for each task.
All filters are summarized into a boxplot whereas using "no filter" appears as a standalone observation.

```{r filter-effect-all-vs-borda-filter, warning=FALSE, dev = c("png", "pdf")}
results_aggr2 <- df_perf %>%
  na.omit() %>% 
  filter(learner_group != "Ridge-MBO", learner_group != "Ridge-CV") %>%
  filter(learner_group != "Lasso-MBO", learner_group != "Lasso-CV") %>%
  mutate(learner_group = recode_factor(learner_group, `XGBoost` = "XG")) %>%
  group_by(learner_group, task.id, filter) %>% 
  # we actually took the mean already in chunk 'aggr-perf'. This is only to get
  # summarise() working
  summarise(perf = mean(rmse_aggr))

results_aggr2 %>%
  ggplot(aes(x = perf, y = learner_group)) +
  geom_beeswarm(
    data = results_aggr2[results_aggr2$filter != "Borda", ],
    shape = 3, size = 1.1, aes(color = "Simple Filter"),
    groupOnX = FALSE
  ) +
  geom_point(data = results_aggr2[results_aggr2$filter == "Borda", ], 
             shape = 19, size = 2, aes(color = "Borda Filter")) +
  facet_wrap(~task.id) +
  scale_color_nejm() +
  labs(x = "RMSE", y = "Task", colour = NULL) +
  guides(size = FALSE) +
  scale_x_continuous(limits = c(30, 48)) +
  theme_pubr() +
  theme(
    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
    axis.title.y = element_blank(),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
  )
```
